\documentclass[table]{gtech}
\PassOptionsToPackage{table,usenames,dvipsnames}{xcolor}
\usepackage{xcolor}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{comment}
\usepackage{bigdelim}
\usepackage{longtable}
\usepackage{tabularray}
\usepackage{wrapfig}
\usepackage[most]{tcolorbox}
\usepackage{url}
\usepackage{float}
\usepackage{datatool}
\usepackage{enumitem}
\usepackage{subcaption} 
\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
\usepackage{pifont}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage[justification=centering]{caption}
\let\comment\undefined
\RequirePackage{tgpagella} 
\RequirePackage{mathpazo}  
\RequirePackage{inconsolata}
\usepackage{makecell}
\usepackage{tcolorbox}

\renewcommand{\title}[1]{\newcommand{\titlelist}{{\huge\fontfamily{optimistic}\selectfont #1}}}

\newcommand{\model}{\texttt{Ring-1T}}
\definecolor{prompt}{HTML}{5f84e4}
\definecolor{img}{HTML}{820100}
\newcommand{\best}[1]{{\textbf{#1}}}
\newcommand{\second}[1]{{\underline{#1}}}

\newcommand\myboxed[1]{{{%
    \setlength\fboxsep{8pt}%
     \boxed{#1}}}}


\newcommand{\cmark}{\textcolor{green}{\ding{51}}}    
\newcommand{\xmark}{\textcolor{red}{\ding{55}}}     
\usepackage{arydshln}
\definecolor{CQColor}{rgb}{0.0,0.0,1.0} 
\definecolor{TABLE_LINE}{rgb}{0.88,0.96,1.0}

\usepackage{colortbl}
\usepackage{amssymb}
\usepackage{pifont}
\usepackage{booktabs,multirow}
\usepackage{makecell}
\usepackage{tabulary}
\usepackage{fontawesome5}
\usepackage{bbding}
\usepackage{multicol}
\newcommand \blfootnote[1]{
    \begingroup
        \renewcommand
        \thefootnote{}\footnote{#1}
        \addtocounter{footnote}{-1}
        \vspace{-1ex}
    \endgroup
}
\newlength\savewidth\newcommand\shline{\noalign{\global\savewidth\arrayrulewidth
  \global\arrayrulewidth 1pt}\hline\noalign{\global\arrayrulewidth\savewidth}}
\newcommand\hshline{\noalign{\global\savewidth\arrayrulewidth
  \global\arrayrulewidth 0.6pt}\hline\noalign{\global\arrayrulewidth\savewidth}}
\newcommand{\tablestyle}[2]{\setlength{\tabcolsep}{#1}\renewcommand{\arraystretch}{#2}\centering\footnotesize}
\newcommand{\placeholder}[2]{%
  \noindent
  \fbox{%
    \rule{0pt}{#2}%
    \makebox[#1][c]{}
  }%
}

\newcommand{\jia}[1]{\textcolor{red}{[Jia: #1]}}
\newcommand{\zy}[1]{\textcolor{purple}{[ZY: #1]}}
\newcommand{\TODO}[1]{\textcolor{blue}{[TODO: #1]}}


\title{\textcolor[HTML]{0369ff}{Every Step Evolves}: Scaling Reinforcement Learning for Trillion-Scale Thinking Model}

\author[*]{Ling Team, Inclusion AI}

\contribution[*]{See Contributions section (Sec.~\ref{sec:contri}) for full author list.}

\abstract{\fontsize{11pt}{12pt} 
We present \model{}, the first open-source, state-of-the-art thinking model with a trillion-scale parameter. It features 1 trillion total parameters and activates approximately 50 billion per token. Training such models at a trillion-parameter scale introduces unprecedented challenges, including train-inference misalignment, inefficiencies in rollout processing, and bottlenecks in the RL system. To address these, we pioneer three interconnected innovations: (1) IcePop stabilizes RL training via token-level discrepancy masking and clipping, resolving instability from training-inference mismatches; (2) C3PO++ improves resource utilization for long rollouts under a token budget by dynamically partitioning them, thereby obtaining high time efficiency; and (3) ASystem, a high-performance RL framework designed to overcome the systemic bottlenecks that impede trillion-parameter model training. \model{} delivers breakthrough results across critical benchmarks: 93.4 on AIME-2025, 86.72 on HMMT-2025, 2088 on CodeForces, and 55.94 on ARC-AGI-1. Notably, it attains a silver medal-level result on the IMO-2025, underscoring its exceptional reasoning capabilities. By releasing the complete 1T parameter MoE model to the community, we provide the research community with direct access to cutting-edge reasoning capabilities. This contribution marks a significant milestone in democratizing large-scale reasoning intelligence and establishes a new baseline for open-source model performance. 
}
\date{Oct 22, 2025\vspace{-1mm}}
\gtechdata[Code]{\url{https://github.com/inclusionAI/Ring-V2}\vspace{-1mm}}
\gtechdata[Model]{\url{https://huggingface.co/inclusionAI/Ring-1T}}

\begin{document}
\maketitle

\input{sections/intro}
\input{sections/method/method}
\input{sections/eval/eval}
\input{sections/conclu}
\input{sections/author}

\bibliographystyle{assets/plainnat}
\bibliography{main}

\clearpage
\beginappendix
\input{sections/appendix}

\end{document}